Sampling requirements to detect ecosystem change

### 2020-02-27 R/manuscript_draft.Rmd
added comments to figure and discussion

We began with a theoretical exploration of the sampling requirements to detect a changepoint. We modeled a simple first order autoregressive (AR-1) process (the discrete-time version of the Ornstein–Uhlenbeck process) with a response variable that represents either population size, biodiversity, or some metric of community composition. The model includes temporal autocorrelation ($\phi$), the mean of the process ($\mu_X$), and a white noise term ($w_t$). The white noise term is a normal distribution with mean ($\mu_w$) and variance ($\sigma$).

$$
\begin{array}{rcl} 
X_t & = & \mu_X + \phi (X_{t-1} - \mu_X) + w_t \\ 
w_t & \sim & Norm(\mu_w,\sigma)
\end{array}
$$
We included a changepoint by shifting $\mu_X$ at time $\tau$ given a specific shift size ($\delta$).

@@

```{r theoretical_results,message=F,warning=F,echo=F,eval=F}

@@

```{r simulation_min_time_plot, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. The default parameters are $\\sigma = 3$, $\\phi = 0.5$, and a shift size $=0.75$. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. \\label{fig:simulation_min_time}',message=F,warning=F,echo=F,eval=T}

>>>>

We began with a theoretical exploration of the sampling requirements to detect a changepoint. We modeled a simple first order autoregressive (AR-1) process (the discrete-time version of the Ornstein–Uhlenbeck process) with a response variable ($X_t$) that represents either population size, biodiversity, or some metric of community composition at time $t$. The model includes temporal autocorrelation ($\phi$), the mean of the process ($\mu_X$), and a white noise term ($w_t$). The white noise term is a normal distribution with mean ($\mu_w$) and variance ($\sigma^2$):

$$
\begin{array}{rcl} 
X_t & = & \mu_X + \phi (X_{t-1} - \mu_X) + w_t \\ 
w_t & \sim & Norm(\mu_w,\sigma^2).
\end{array}
$$
We included a changepoint by shifting $\mu_X$ at time $\tau$ given a specific shift size ($\delta$).

@@

```{r theoretical_results_old,message=F,warning=F,echo=F,eval=F}

@@

```{r simulation_min_time_plot, fig.cap='Minimum number of samples required for 0.8 statistical power given different levels of (a) temporal variability ($\\sigma$), (b) temporal autocorrelation ($\\phi$), and (c) shift size. Random sampling was used with the default parameters are $\\sigma = 3$, $\\phi = 0.5$, and a shift size $=0.75$. The exact timing of the true changepoint varied for each simulation between time steps 30 and 70. \\label{fig:simulation_min_time}',message=F,warning=F,echo=F,eval=T}
